Baseline centering problem

### 2018-12-12 paper/tangotd.tex

>>>>

\paragraph{Double sampling problem: using a model for one of the
samples.} Double sampling allows us to minimize the $L^2$ norm of the
Bellman gap. This is only possible if we have model of the environment,
that allows for sampling $s'$ given $s$ and $a$.

However, using Bellman on a model is dangerous because we compute the
value function of the model, and even if the model is accurate for one
step $s\to s'$, it may quickly diverge when iterated. Therefore its value
function may be quite different from the true value function. The model
needs to be accurate at time horizon $1/(1-\gamma)$.

But in double-sampling, one could also use one real transition and one
simulated transition. We show here that this is equivalent, at first order,
to learning the value function of an environment in which every step is
either a real or simulated step with probability $1/2$. This should be a
bit more stable, as the real transitions may keep the model from
hallucinating completely irrealistic trajectories.

The $L^2(\rho)$ norm of the Bellman gap is $\transp{(\Delta
V-R)}\rho(\Delta V-R)$. Its gradient with respect to $V$ is
$2\transp{\Phi}\transp{\Delta}\rho (\Delta V-r)$. The presence of
$\transp{\Delta}\rho\Delta$ indicates double sampling; this operator is
not samplable.

Let us assume that we have a model $\tilde P$ of the operator $P$. Thus
we can sample $\transp{\tilde \Delta}\rho \Delta$ (just sample one
transition $s\to s'$ from $\rho P$, and a second transition $s\to \tilde
s'$ from the same $s$). We can also sample from $\transp{\tilde
\Delta}\rho R$. Moreover, \emph{if rewards during $s\to s'$ are
independent from $s'$}, we can sample from $\transp{\Delta}\rho R$. From
now on, I will assume that the reward collected during a transition $s\to
s'$ depends only on $s$. \footnote{This means the state $s$ is a
sufficient statistic. If the reward depends only on $s'$ instead, we can
just
shift the rewards by one step and pretend they are collected at the next
step; therefore we are back to a reward that depends only on $s$. The
only case not covered is if the reward really depends on the pair
$(s,s')$; in that case, the state $s'$ is not quite a sufficient
statistic of the process.

In the latter case, one would also need a model $\tilde R$ of rewards,
and then we could sample from $\transp{\tilde \Delta}\rho R$ and from
$\transp{\Delta}\rho\tilde R$.}

Let us consider the norm
\begin{equation}
\norm{\Delta V-R}^2_{P,\tilde P}\deq \transp{(\tfrac12 (\Delta+\tilde
\Delta)
V-R)}\rho(\tfrac12 (\Delta+\tilde \Delta)V-R)
\end{equation}
which is the $L^2(\rho)$-norm of the Bellman gap of an environment
$\frac12 (P+\tilde P)$ in which half of the transitions are real or
simulated.

The gradient of this norm is
\begin{equation}
\tfrac12 \transp{\Phi}\transp{(\Delta+\tilde
\Delta)
}\rho((\Delta+\tilde \Delta)V-R)
\end{equation}

We can easily compute
\begin{equation}
\tfrac12 \transp{\Phi}\transp{\Delta}
\rho(\tilde \Delta V-R)
+
\tfrac12 \transp{\Phi}\transp{\tilde \Delta}
\rho(\Delta V-R)
\end{equation}
